Manuscript

### 2025-03-22 Manuscript_MD_ScoreReLiability.qmd
add methods to second analysis

@tbl-tauES showed that for all phenomena, where we identified
statistically significant ES heterogeneity, the heterogeneity was even
larger after the attenuation correction procedure. Subsequently, in
Appendix A we demonstrate that if the $R_2$ metric is larger than
$R_1^3$, ES heterogeneity is bound to be larger after applying said
procedure. @tbl-R1R2 reports the meta-analytic estimates of observed and
true score variances, including their heterogeneity and metrics $R_1$
and $R_2$ across all 12 phenomena. In @tbl-R1R2 we see that for ten out
of 12 phenomena, the $R_2$ metric is larger than the $R_1$ metric.
Additionally, we observe that across nine phenomena, $R_2$ is larger
than .96. This implies that of the heterogeneity observed in score
variances, 97% or more could not be explained by differences in random
error introduced by the measuring instrument, but by genuine differences
in the underlying distribution.

>>>>

@tbl-tauES showed that for all phenomena, where we identified
statistically significant ES heterogeneity, the heterogeneity was even
larger after the attenuation correction procedure. Subsequently, in
Appendix A we demonstrate that if the $R_2$ metric is larger than
$R_1^3$, ES heterogeneity is bound to be larger after applying said
procedure. Here, we will re-analyse the data, estimating $R_1$ and $R_2$
for each phenomenon respectively. This allows us to assess in how far
the estimates of ES heterogeneity align with our claims regarding the
$R_1$ and $R_2$ metrics.

#### Methods

Estimation of $R_1$ and $R_2$ require estimates of meta-analytic mean
and heterogeneity of both the total score variance $\sigma^2_X$ and the
true score variance $\sigma^2_T$. Elsewhere, we proposed the method
boot-err (ref.) as an alternative method to test for and estimate the
differences in error score variance $\sigma^2_E$. In this text, we
deliver detailed instructions on how estimates for mean and
heterogeneity of total score variance, and the score variance components
can be derived. For the total score variance $\sigma^2_X$, all we need
to do is to appropriately transform the observed estimates using the
variance-stabilizing properties of the natural log. Subsequently, we use
bootstrapping to derive robust standard errors for these transformed
observations. Using a random-efffects meta-analysis, we derive estimates
of mean and heterogeneity, which need to be backtransformed to the
original scale. Appropriate equation to back-transform estimates can be
found in Appendix B.

For the true score variance $\sigma^2_T$, we found that directly
estimating its heterogeneity, even if an appropriate
variance-stabilizing transformation is used, is severely biased.
Instead, we can estimate the error score variance $\sigma^2_E$ using an
estimate of score reliability. Performing the boot-err, which, again,
implies transformations using the natural log, deriving standard errors
using Bootstrapping, running a random-effects meta-analysis and
back-transforming the estimates, nets us estimates of mean and
heterogeneity of $\sigma^2_E$. Since we already have the same estimates
available for the total score variance, we can estimate meta-analytic
mean and heterogeneity of true score variance.

$$\hat{\mu}_{\sigma^2_T} = \hat{\mu}_{\sigma^2_X} - \hat{\mu}_{\sigma^2_E} \ \ \ \ \ \ and \ \ \ \ \  \hat{\tau}^2_{\sigma^2_T} = \hat{\tau}^2_{\sigma^2_X} - \hat{\tau}^2_{\sigma^2_E}$$ {#eq-meantauT}

@eq-meantauT provides estimates of meta-analytic mean of true score
variance and heterogeneity. These estimates are largely unbiased, unlike
estimates stemming from using the boot-err to directly run meta-analytic
models on true score variance estimates. Interested readers are referred
to Beinhauer et al. (ref.)

#### Results

@tbl-R1R2 reports the meta-analytic estimates of observed and true score
variances, including their heterogeneity and metrics $R_1$ and $R_2$
across all 12 phenomena. In @tbl-R1R2 we see that for ten out of 12
phenomena, the $R_2$ metric is larger than the $R_1$ metric.
Additionally, we observe that across nine phenomena, $R_2$ is larger
than .96. This implies that of the heterogeneity observed in score
variances, 97% or more could not be explained by differences in random
error introduced by the measuring instrument, but by genuine differences
in the underlying distribution.
